---
title: "HW3 Step"
author: "Dan Fanelli"
date: "July 2, 2016"
output: html_document
---

### What we handed in: Normalized Backward Selection

model3a <- glm(target ~ nox + age + rad + tax + ptratio + black + medv, data = training_norm, family = binomial)

### What we handed in: Normalized Forward Selection

model3b <- glm(target ~ nox + rad + tax + ptratio + black + medv + age + dis + zn + lstat, data = data3b, family = binomial)

# Functions to do step and stepAIC:

```{r}
library(MASS)
data.all <- read.csv('crime-training-data.csv')

do_step_aic <- function(the_direction){
  fit <- lm(target ~ ., data=data.all)
  summary(stepAIC(fit, direction=the_direction, trace = FALSE))
}

do_step_by_step <- function(the_direction){
  fit <- lm(target ~ ., data=data.all)
  fit <- step(fit, direction=the_direction, trace = 0)
  summary(fit)
}
```

### All 6 Combos yield: (nox, age, rad, medv)

# Forwards Comparison

### Forward: stepAIC() 

```{r}
do_step_aic("forward")
```

### Forward: step()

```{r}
do_step_by_step("forward")
```

# Backward Comparison

### Backward: stepAIC() 

```{r}
do_step_aic("backward")
```

### Backward: step()

```{r}
do_step_by_step("backward")
```

# Both Comparison

### Both: stepAIC() 

```{r}
do_step_aic("both")
```

### Both: step()

```{r}
do_step_by_step("both")
```


