---
title: "HW5_YQ"
author: "Youqing Xiang"
date: "July 16, 2016"
output: pdf_document
---

# Data Exploration

```{r,echo=FALSE, warning=FALSE, message=FALSE}
library(ggplot2)
library(MASS)
library(knitr)
library(gridExtra)
library(grid)
library(dplyr)
library(Amelia)

wine <- read.csv("wine-training-data.csv")
wineN <- select(wine,-c(INDEX))
df <- data.frame(Variable = colnames(wineN), NAs =
                        c(0,0,0,0,616,638,647,682,0,395,1210,653,0,0,3359))
df$Percent_NAs <- round(df$NAs/dim(wineN)[1],3)
kable(df)
```

This dataset includes 12795 observations with 15 variables (`Index` columns excluded) in total. As the table shows above, there are a fair number of NAs and `STARS` variable has the most NAs.

```{r,echo=FALSE, warning=FALSE, message=FALSE}
missmap(wine)
```

The above matrix of the NAs doesn't show any pattern. So, we could consider to replace the NA values with certain method to avoid potentially losing a lardge amount to data if we just simply drop off the NA values during data preparation process.

Since `STARS` variable has the most NAs, which accounts for 26.3% of data points, it is worthwhile to check how `STARS` variable NAs affect other variables, especially the `TARGET` variable. Here I created a new varialbe: `STARS_Present`, which is a categorical variable, equals to 0 when `STARS` value is missing and 1 when `STARS` value is present. And then I show density plots for each variable grouped by `STARS_Present`.

```{r,echo=FALSE, warning=FALSE, message=FALSE}
wine$STARS_Present <- factor(ifelse(is.na(wine$STARS),0,1))

Target_df <- select(wine,c(TARGET,STARS_Present))
Target_p1 <- ggplot(Target_df, aes(x=TARGET)) + geom_density(aes(colour=STARS_Present)) +
  theme(legend.position=c(0.6,0.8))

FixedAcidity_df <- select(wine,c(FixedAcidity,STARS_Present))
FixedAcidity_p1 <- ggplot(FixedAcidity_df, aes(x=FixedAcidity)) + 
  geom_density(aes(colour=STARS_Present)) +
  theme(legend.position=c(0.75,0.8))

VolatileAcidity_df <- select(wine,c(VolatileAcidity,STARS_Present))
VolatileAcidity_p1 <- ggplot(VolatileAcidity_df, aes(x=VolatileAcidity)) + 
  geom_density(aes(colour=STARS_Present)) +
  theme(legend.position=c(0.75,0.8))

CitricAcid_df <- select(wine,c(CitricAcid,STARS_Present))
CitricAcid_p1 <- ggplot(CitricAcid_df, aes(x=CitricAcid)) + 
  geom_density(aes(colour=STARS_Present)) +
  theme(legend.position=c(0.75,0.8))

ResidualSugar_df <- select(wine,c(ResidualSugar,STARS_Present))
ResidualSugar_p1 <- ggplot(ResidualSugar_df, aes(x=ResidualSugar)) + 
  geom_density(aes(colour=STARS_Present)) +
  theme(legend.position=c(0.75,0.8))

Chlorides_df <- select(wine,c(Chlorides,STARS_Present))
Chlorides_p1 <- ggplot(Chlorides_df, aes(x=Chlorides)) + 
  geom_density(aes(colour=STARS_Present)) +
  theme(legend.position=c(0.75,0.8))

FreeSulfurDioxide_df <- select(wine,c(FreeSulfurDioxide,STARS_Present))
FreeSulfurDioxide_p1 <- ggplot(FreeSulfurDioxide_df, aes(x=FreeSulfurDioxide)) + 
  geom_density(aes(colour=STARS_Present)) +
  theme(legend.position=c(0.75,0.8))

TotalSulfurDioxide_df <- select(wine,c(TotalSulfurDioxide,STARS_Present))
TotalSulfurDioxide_p1 <- ggplot(TotalSulfurDioxide_df, aes(x=TotalSulfurDioxide)) + 
  geom_density(aes(colour=STARS_Present)) +
  theme(legend.position=c(0.75,0.8))

Density_df <- select(wine,c(Density,STARS_Present))
Density_p1 <- ggplot(Density_df, aes(x=Density)) + 
  geom_density(aes(colour=STARS_Present)) +
  theme(legend.position=c(0.75,0.8))

pH_df <- select(wine,c(pH,STARS_Present))
pH_p1 <- ggplot(pH_df, aes(x=pH)) + 
  geom_density(aes(colour=STARS_Present)) +
  theme(legend.position=c(0.75,0.8))

Sulphates_df <- select(wine,c(Sulphates,STARS_Present))
Sulphates_p1 <- ggplot(Sulphates_df, aes(x=Sulphates)) + 
  geom_density(aes(colour=STARS_Present)) +
  theme(legend.position=c(0.75,0.8))

Alcohol_df <- select(wine,c(Alcohol,STARS_Present))
Alcohol_p1 <- ggplot(Alcohol_df, aes(x=Alcohol)) + 
  geom_density(aes(colour=STARS_Present)) +
  theme(legend.position=c(0.75,0.8))

LabelAppeal_df <- select(wine,c(LabelAppeal,STARS_Present))
LabelAppeal_p1 <- ggplot(LabelAppeal_df, aes(x=LabelAppeal)) + 
  geom_density(aes(colour=STARS_Present)) +
  theme(legend.position=c(0.75,0.8))

AcidIndex_df <- select(wine,c(AcidIndex,STARS_Present))
AcidIndex_p1 <- ggplot(AcidIndex_df, aes(x=AcidIndex)) + 
  geom_density(aes(colour=STARS_Present)) +
  theme(legend.position=c(0.75,0.8))

AcidIndex_df <- select(wine,c(AcidIndex,STARS_Present))
AcidIndex_p1 <- ggplot(AcidIndex_df, aes(x=AcidIndex)) + 
  geom_density(aes(colour=STARS_Present)) +
  theme(legend.position=c(0.75,0.8))

grid.arrange(FixedAcidity_p1,VolatileAcidity_p1,ncol=2,nrow=1)
grid.arrange(CitricAcid_p1,ResidualSugar_p1,ncol=2,nrow=1)
grid.arrange(Chlorides_p1,FreeSulfurDioxide_p1,ncol=2,nrow=1)
grid.arrange(TotalSulfurDioxide_p1,Density_p1,ncol=2,nrow=1)
grid.arrange(pH_p1,Sulphates_p1,ncol=2,nrow=1)
grid.arrange(Alcohol_p1,LabelAppeal_p1,ncol=2,nrow=1)
grid.arrange(AcidIndex_p1,Target_p1,ncol=2,nrow=1)
```

From the above density plots, we can see that `STARS` variable NAs have big effect on `TARGET` distribution but no obvious effects on other variables. And we can also conclude that **STARS variable NAs** is actually predictive of the target. Overall, this part of analysis suggests us that we should treat **STARS variable NAs** as category variable. 