---
title: "hw4_cjf"
author: "Chris Fenton"
date: "July 4, 2016"
output: html_document
---
Libraries:
```{r}
library(stringr)
library(PerformanceAnalytics)
library(aod)
library(ggplot2)
library(Rcpp)
library(Amelia)
```

Loading the data:

```{r}
ins <- read.csv('insurance_training_data.csv',na.strings=c("","NA"),stringsAsFactors = FALSE)
```

# Data Exploration



```{r}
#check for missing values
sapply(ins, function(x) sum(is.na(x)))

```
There are 2116 incomplete observations (missing at least one variable value), these will need to either be imputed or dropped.

```{r}

#chart.Correlation(ins[,-c(9,11,12,13,14,16,19,20,23,26)])
```
# Data Transformation


Converting the money character fields to numeric.
```{r}
blue_book <- unname(sapply(ins$BLUEBOOK, str_replace_all, '[,$]', ''))
blue_book <- as.numeric(blue_book)

income <- unname(sapply(ins$INCOME, str_replace_all, '[,$]', ''))
income <- as.numeric(income)

home_val <- unname(sapply(ins$HOME_VAL, str_replace_all, '[,$]', ''))
home_val <- as.numeric(home_val)

old_claim <- unname(sapply(ins$OLDCLAIM, str_replace_all, '[,$]', ''))
old_claim <- as.numeric(old_claim)

ins$BLUEBOOK <- blue_book
ins$INCOME <- income
ins$HOME_VAL <- home_val
ins$OLDCLAIM <- old_claim
```

Converting the categoricals to factors:
```{r}
ins$CAR_TYPE <- factor(ins$CAR_TYPE)
ins$CAR_USE <- factor(ins$CAR_USE)
ins$EDUCATION <- factor(ins$EDUCATION)
ins$JOB <- factor(ins$JOB)
ins$MSTATUS <- factor(ins$MSTATUS)
ins$PARENT1 <- factor(ins$PARENT1)
ins$RED_CAR <- factor(ins$RED_CAR)
ins$REVOKED <- factor(ins$REVOKED)
ins$SEX <- factor(ins$SEX)
ins$URBANICITY <- factor(ins$URBANICITY)
```

Create numeric data frame:
```{r}
ins_num <- ins[,-c(1,9,11,12,13,14,16,19,20,23,26)]
```



# Multiple Linear Regression Model to Predict TARGET_AMT

Only look at observations that paid out.
```{r}
ins_payout <- subset(ins, TARGET_AMT != 0)
```

Adding log of Target Amount
```{r}
ins_payout$TARGET_AMT_LOG <- log(ins_payout$TARGET_AMT)
ins_payout$BLUEBOOK_LOG <- log(ins_payout$BLUEBOOK)
ins_payout$INCOME_LOG <- log(ins_payout$INCOME)

ins_payout$INCOME_LOG <- log(ins_payout$INCOME)

```


Check to confirm that they were all in crashes.
```{r}
ins_payout2 <- ins[ins$TARGET_AMT != 0 & ins$TARGET_FLAG == 1,]
```

Look to see what variables have linear relationship with TARGET_AMT, thus eligible to be used/transformed into a regression model.
```{r}
plot(ins_payout$YOJ, ins_payout$TARGET_AMT)

```


```{r}
blue_book_model_log <- lm(TARGET_AMT_LOG ~ BLUEBOOK_LOG, data=ins_payout)
blue_book_model_2 <- lm(TARGET_AMT_LOG ~ BLUEBOOK_LOG + CAR_AGE + CAR_TYPE, data=ins_payout)

summary(blue_book_model_log)
summary(blue_book_model_2)
```

The variables `CAR_AGE` and `CAR_TYPE` were indidcated as "probably effect[ing] the payout if there is a crash" in the dataset description. However, they are not found to be signficantly impacting the `TARGET_AMT` variable. Furthermore, the adjusted R-Squared is marginally higher without these variables. Thus `blue_book_model_log` is the better model to proceed with. 

Verifying the assumptions of `blue_book_model_log`:

Linearity:

```{r}
plot(ins_payout$BLUEBOOK_LOG, ins_payout$TARGET_AMT_LOG)

cor(ins_payout$BLUEBOOK_LOG, ins_payout$INCOME_LOG)

```

Normality, homoscedasticity, and independence of residuals:

```{r}
plot(blue_book_model_log$residuals)
```