---
title: "HW4_YQ"
author: "Youqing Xiang"
date: "July 7, 2016"
output: pdf_document
---
# Data Exploration

```{r}
library(PerformanceAnalytics)
library(ggplot2)
library(gridExtra)
library(knitr)
library(lattice)
library(caret)
library(tidyr)
library(dplyr)
library(pROC)
library(car)

data <- read.csv('insurance_training_data.csv')
summary(data)
dim(data)
```

# Data Preparation
## Deal with Missing Data and Nonsense Data
```{r}
# AGE
data <- data[!is.na(data$AGE),]

# YOJ
dataN <- data[!is.na(data$YOJ),]

# CAR_AGE
dataN <- dataN[!is.na(dataN$CAR_AGE),]
dataN <- dataN[dataN$CAR_AGE >= 0,]
```

## Data Transformation
### KIDSDRIV

```{r}
# Before transformation
p1 <- ggplot(dataN, aes(KIDSDRIV)) + geom_histogram()
p2 <- ggplot(dataN, aes(factor(TARGET_FLAG), KIDSDRIV)) + geom_boxplot()
p3 <- ggplot(dataN, aes(factor(KIDSDRIV), TARGET_AMT)) + geom_boxplot()
grid.arrange(p1,p2,p3,ncol=3,nrow=1)

# Data transformation
dataN$KIDSDRIV_N <- ifelse(dataN$KIDSDRIV == 0, 0, 1)
dataN$KIDSDRIV_N <- as.factor(dataN$KIDSDRIV_N)

# After transformation
t <- as.data.frame(table(KIDSDRIV_N=dataN$KIDSDRIV_N, TARGET_FLAG=dataN$TARGET_FLAG))
p1 <- ggplot(t, aes(factor(TARGET_FLAG), Freq)) + geom_bar(aes(fill=KIDSDRIV_N),stat='identity',position=position_dodge())
p2 <- ggplot(dataN, aes(factor(KIDSDRIV_N), TARGET_AMT)) + geom_boxplot()
grid.arrange(p1,p2,ncol=2,nrow=1)
```

### HOMEKIDS

```{r}
# Before transformation
p1 <- ggplot(dataN, aes(HOMEKIDS)) + geom_histogram()
p2 <- ggplot(dataN, aes(factor(TARGET_FLAG), HOMEKIDS)) + geom_boxplot()
p3 <- ggplot(dataN, aes(factor(HOMEKIDS), TARGET_AMT)) + geom_boxplot()
grid.arrange(p1,p2,p3,ncol=3,nrow=1)

# Data transformation
dataN$HOMEKIDS_N <- ifelse(dataN$HOMEKIDS == 0, 0, 1)
dataN$HOMEKIDS_N <- as.factor(dataN$HOMEKIDS_N)

# After transformation
t <- as.data.frame(table(HOMEKIDS_N=dataN$HOMEKIDS_N, TARGET_FLAG=dataN$TARGET_FLAG))
p1 <- ggplot(t, aes(factor(TARGET_FLAG), Freq)) + geom_bar(aes(fill=HOMEKIDS_N),stat='identity', position=position_dodge())
p2 <- ggplot(dataN, aes(factor(HOMEKIDS_N), TARGET_AMT)) + geom_boxplot()
grid.arrange(p1,p2,ncol=2,nrow=1)
```

### INCOME

```{r}
dataN$INCOME <- as.numeric(dataN$INCOME)
```

### HOME_VAL

```{r}
dataN$HOME_VAL <- as.numeric(dataN$HOME_VAL)
```

### EDUCATION

```{r}
# Before transformation
t <- as.data.frame(table(EDUCATION=dataN$EDUCATION, TARGET_FLAG=dataN$TARGET_FLAG))
p1 <- ggplot(t, aes(factor(TARGET_FLAG), Freq)) + geom_bar(aes(fill=EDUCATION),stat='identity',
                                                           position=position_dodge())
p2 <- ggplot(dataN, aes(factor(EDUCATION), TARGET_AMT)) + geom_boxplot()
grid.arrange(p1,p2,ncol=2,nrow=1)

# Data transformation
dataN$EDUCATION_N <- ifelse(dataN$EDUCATION %in% c('<High School','z_High School'), 0, 1)
dataN$EDUCATION_N <- as.factor(dataN$EDUCATION_N)

# After transformation
t <- as.data.frame(table(EDUCATION_N=dataN$EDUCATION_N, TARGET_FLAG=dataN$TARGET_FLAG))
p1 <- ggplot(t, aes(factor(TARGET_FLAG), Freq)) + geom_bar(aes(fill=EDUCATION_N),stat='identity',
                                                           position=position_dodge())
p2 <- ggplot(dataN, aes(factor(EDUCATION_N), TARGET_AMT)) + geom_boxplot()
grid.arrange(p1,p2,ncol=2,nrow=1)
```

### JOB

```{r}
# Before transformation
t <- as.data.frame(table(JOB=dataN$JOB, TARGET_FLAG=dataN$TARGET_FLAG))
p1 <- ggplot(t, aes(factor(TARGET_FLAG), Freq)) + geom_bar(aes(fill=JOB),stat='identity',
                                                           position=position_dodge())
p2 <- ggplot(dataN, aes(factor(JOB), TARGET_AMT)) + geom_boxplot()
grid.arrange(p1,p2,ncol=2,nrow=1)

# Data transformation
dataN$JOB_N <- ifelse(dataN$JOB %in% c('Doctor','Lawyer','Manager','Professional'), 1, 0)
dataN$JOB_N <- as.factor(dataN$JOB_N)

# After transformation
t <- as.data.frame(table(JOB_N=dataN$JOB_N, TARGET_FLAG=dataN$TARGET_FLAG))
p1 <- ggplot(t, aes(factor(TARGET_FLAG), Freq)) + geom_bar(aes(fill=JOB_N),stat='identity',
                                                           position=position_dodge())
p2 <- ggplot(dataN, aes(factor(JOB_N), TARGET_AMT)) + geom_boxplot()
grid.arrange(p1,p2,ncol=2,nrow=1)
```

### BLUEBOOK

```{r}
dataN$BLUEBOOK <- as.numeric(dataN$BLUEBOOK)
```

### TIF

```{r}
# Before transformation
p1 <-  ggplot(dataN, aes(TIF)) + geom_histogram()
p2 <- ggplot(dataN, aes(factor(TARGET_FLAG), TIF)) + geom_boxplot()
p3 <- ggplot(dataN, aes(TIF, TARGET_AMT)) + geom_point()
grid.arrange(p1,p2,p3,ncol=3,nrow=1)

ggplot(dataN, aes(x=TIF)) + geom_density(aes(colour=factor(TARGET_FLAG)))

# Data transformation
dataN$TIF_N <- ifelse(dataN$TIF > 1, 1, 0)
dataN$TIF_N <- as.factor(dataN$TIF_N)

# After transformation
t <- as.data.frame(table(TIF_N=dataN$TIF_N, TARGET_FLAG=dataN$TARGET_FLAG))
p1 <- ggplot(t, aes(factor(TARGET_FLAG), Freq)) + geom_bar(aes(fill=TIF_N),stat='identity',
                                                           position=position_dodge())
p2 <- ggplot(dataN, aes(factor(TIF_N), TARGET_AMT)) + geom_boxplot()
grid.arrange(p1,p2,ncol=2,nrow=1)
```

### CAR_TYPE

```{r}
# Before transformation
t <- as.data.frame(table(CAR_TYPE=dataN$CAR_TYPE, TARGET_FLAG=dataN$TARGET_FLAG))
p1 <- ggplot(t, aes(factor(TARGET_FLAG), Freq)) + geom_bar(aes(fill=CAR_TYPE),stat='identity',
                                                           position=position_dodge())
p2 <- ggplot(dataN, aes(factor(CAR_TYPE), TARGET_AMT)) + geom_boxplot()
grid.arrange(p1,p2,ncol=2,nrow=1)

# Data transformation
dataN$CAR_TYPE_N <- ifelse(dataN$CAR_TYPE == 'Minivan', 0, 1)
dataN$CAR_TYPE_N <- as.factor(dataN$CAR_TYPE_N)

# After transformation
t <- as.data.frame(table(CAR_TYPE_N=dataN$CAR_TYPE_N, TARGET_FLAG=dataN$TARGET_FLAG))
p1 <- ggplot(t, aes(factor(TARGET_FLAG), Freq)) + geom_bar(aes(fill=CAR_TYPE_N),stat='identity',
                                                           position=position_dodge())
p2 <- ggplot(dataN, aes(factor(CAR_TYPE_N), TARGET_AMT)) + geom_boxplot()
grid.arrange(p1,p2,ncol=2,nrow=1)
```

### OLDCLAIM

```{r}
# Before transformation
dataN$OLDCLAIM <- as.numeric(dataN$OLDCLAIM)
p1 <-  ggplot(dataN, aes(OLDCLAIM)) + geom_histogram()
p2 <- ggplot(dataN, aes(factor(TARGET_FLAG), OLDCLAIM)) + geom_boxplot()
p3 <- ggplot(dataN, aes(OLDCLAIM, TARGET_AMT)) + geom_point()
grid.arrange(p1,p2,p3,ncol=3,nrow=1)

ggplot(dataN, aes(x=OLDCLAIM)) + geom_density(aes(colour=factor(TARGET_FLAG))) + 
  geom_vline(xintercept = 150)

# Data transformation
dataN$OLDCLAIM_N <- ifelse(dataN$OLDCLAIM > 150, 1, 0)
dataN$OLDCLAIM_N <- as.factor(dataN$OLDCLAIM_N)

# After transformation
t <- as.data.frame(table(OLDCLAIM_N=dataN$OLDCLAIM_N, TARGET_FLAG=dataN$TARGET_FLAG))
p1 <- ggplot(t, aes(factor(TARGET_FLAG), Freq)) + geom_bar(aes(fill=OLDCLAIM_N),stat='identity',position=position_dodge())

p2 <- ggplot(dataN, aes(factor(OLDCLAIM_N), TARGET_AMT)) + geom_boxplot()
grid.arrange(p1,p2,ncol=2,nrow=1)
```

### CLM_FREQ

```{r}
# Before transformation
t <- as.data.frame(table(CLM_FREQ=dataN$CLM_FREQ, TARGET_FLAG=dataN$TARGET_FLAG))
p1 <- ggplot(t, aes(factor(TARGET_FLAG), Freq)) + geom_bar(aes(fill=CLM_FREQ),stat='identity',
                                                           position=position_dodge())
p2 <- ggplot(dataN, aes(factor(CLM_FREQ), TARGET_AMT)) + geom_boxplot()
grid.arrange(p1,p2,ncol=2,nrow=1)

# Data transformation
dataN$CLM_FREQ_N <- ifelse(dataN$CLM_FREQ == 0, 0, 1)
dataN$CLM_FREQ_N <- as.factor(dataN$CLM_FREQ_N)

# After transformation
t <- as.data.frame(table(CLM_FREQ_N=dataN$CLM_FREQ_N, TARGET_FLAG=dataN$TARGET_FLAG))
p1 <- ggplot(t, aes(factor(TARGET_FLAG), Freq)) + geom_bar(aes(fill=CLM_FREQ_N),stat='identity',position=position_dodge())
p2 <- ggplot(dataN, aes(factor(CLM_FREQ_N), TARGET_AMT)) + geom_boxplot()
grid.arrange(p1,p2,ncol=2,nrow=1)
```

### MVR_PTS

```{r}
# Before transformation
t <- as.data.frame(table(MVR_PTS=dataN$MVR_PTS, TARGET_FLAG=dataN$TARGET_FLAG))
p1 <- ggplot(t, aes(factor(TARGET_FLAG), Freq)) + geom_bar(aes(fill=MVR_PTS),stat='identity',
                                                           position=position_dodge())
p2 <- ggplot(dataN, aes(factor(MVR_PTS), TARGET_AMT)) + geom_boxplot()
grid.arrange(p1,p2,ncol=2,nrow=1)

# Data transformation
dataN$MVR_PTS_N <- ifelse(dataN$MVR_PTS %in% c(0,1), 0, 1)
dataN$MVR_PTS_N <- as.factor(dataN$MVR_PTS_N)

# After transformation
t <- as.data.frame(table(MVR_PTS_N=dataN$MVR_PTS_N, TARGET_FLAG=dataN$TARGET_FLAG))
p1 <- ggplot(t, aes(factor(TARGET_FLAG), Freq)) + geom_bar(aes(fill=MVR_PTS_N),stat='identity',
                                                           position=position_dodge())
p2 <- ggplot(dataN, aes(factor(MVR_PTS_N), TARGET_AMT)) + geom_boxplot()
grid.arrange(p1,p2,ncol=2,nrow=1)
```

# Modeling
## TARGET_FLAG
### Splitting data for training and testing models
```{r}
set.seed(45)
inTrain_1 <- createDataPartition(y=dataN$TARGET_FLAG, p=0.7,list=FALSE)
training_1 <- dataN[inTrain_1,]
testing_1 <- dataN[-inTrain_1,]
```

### Model 1 - Using original variables
```{r}
training_1a <- select(training_1, -c(KIDSDRIV_N,HOMEKIDS_N,EDUCATION_N,JOB_N,TIF_N,CAR_TYPE_N,OLDCLAIM_N,CLM_FREQ_N,MVR_PTS_N))
m11 <- glm(TARGET_FLAG ~ . -INDEX-TARGET_AMT, data=training_1a,family = binomial(link='probit'))
#summary(m11)
m12 <- update(m11, .~. -AGE-INCOME-BLUEBOOK-RED_CAR-CAR_AGE-CLM_FREQ-TIF)
#summary(m12)
TARGET_FLAG_m1 <- m12
summary(TARGET_FLAG_m1)
```

### Model 2 - Using transformed variables
```{r}
training_1b <- select(training_1, -c(KIDSDRIV,HOMEKIDS,EDUCATION,JOB,TIF,CAR_TYPE,OLDCLAIM,CLM_FREQ,MVR_PTS))
m21 <- glm(TARGET_FLAG ~ . -INDEX-TARGET_AMT, data=training_1b,family = binomial(link='probit'))
#summary(m21)
m22 <- update(m21, .~. -AGE-INCOME-PARENT1-SEX-BLUEBOOK-RED_CAR-CAR_AGE-OLDCLAIM_N)
#summary(m22)
TARGET_FLAG_m2 <- m22
summary(TARGET_FLAG_m2)
```

### Model 3 - Combining both of original and transformed variables
```{r}
m31 <- glm(TARGET_FLAG ~ KIDSDRIV_N+HOMEKIDS_N+YOJ+HOME_VAL+PARENT1+MSTATUS+SEX+
             EDUCATION_N+JOB_N+TIF_N+CAR_USE+CAR_TYPE_N+OLDCLAIM+REVOKED+MVR_PTS+
             URBANICITY, data=training_1,family = binomial(link='probit'))
#summary(m31)
m32 <- update(m31, .~. -SEX-PARENT1)
#summary(m32)
TARGET_FLAG_m3 <- m32
summary(TARGET_FLAG_m3)
```

## TARGET_AMT
### Splitting data for training and testing models
```{r}
set.seed(1234)
inTrain_2 <- createDataPartition(y=dataN$TARGET_AMT, p=0.7,list=FALSE)
training_2 <- dataN[inTrain_2,]
testing_2 <- dataN[-inTrain_2,]
```

### Model 1 - Using original variables
```{r}
training_2a <- select(training_2, -c(KIDSDRIV_N,HOMEKIDS_N,EDUCATION_N,JOB_N,TIF_N,
                                     CAR_TYPE_N,OLDCLAIM_N,CLM_FREQ_N,MVR_PTS_N))
M11 <- lm( TARGET_AMT~ .-TARGET_FLAG-INDEX, data=training_2a)
#summary(M11)
M12 <- update(M11,.~.-AGE-HOMEKIDS-YOJ-INCOME-SEX-EDUCATION-BLUEBOOK-RED_CAR-OLDCLAIM-CLM_FREQ)
#summary(M12)
TARGET_AMT_m1 <- M12
summary(TARGET_AMT_m1)
```

### Model 2 - Using transformed variables
```{r}
training_2b <- select(training_2, -c(KIDSDRIV,HOMEKIDS,EDUCATION,JOB,TIF,CAR_TYPE,
                                     OLDCLAIM,CLM_FREQ,MVR_PTS))
M21 <- lm( TARGET_AMT~ .-TARGET_FLAG-INDEX, data=training_2b)
#summary(M21)
M22 <- update(M21, .~.-AGE-YOJ-INCOME-PARENT1-SEX-BLUEBOOK-RED_CAR-HOMEKIDS_N-
                EDUCATION_N-OLDCLAIM_N-CLM_FREQ_N)
#summary(M22)
TARGET_AMT_m2 <- M22
summary(TARGET_AMT_m2)
```

### Model 3 - Combining both of original and transformed variables
```{r}
M31 <- lm( TARGET_AMT~HOME_VAL+MSTATUS+TRAVTIME+CAR_USE+REVOKED+CAR_AGE+URBANICITY+
             KIDSDRIV+JOB_N+TIF+CAR_TYPE_N+MVR_PTS+PARENT1, data=training_2)
#summary(M31)
TARGET_AMT_m3 <- M31
summary(TARGET_AMT_m3)
```

# Model Selection
## TARGET_FLAG
### key model statistics measurements

```{r}
# Model1
predict_1 <- predict(TARGET_FLAG_m1, newdata=testing_1, type='response')
glm.pred1 = ifelse(predict_1 > 0.5, 1, 0)
cM1 <- confusionMatrix(glm.pred1, testing_1$TARGET_FLAG, positive = "1")

# Model2
predict_2 <- predict(TARGET_FLAG_m2, newdata=testing_1, type='response')
glm.pred2 = ifelse(predict_2 > 0.5, 1, 0)
cM2 <- confusionMatrix(glm.pred2, testing_1$TARGET_FLAG, positive = "1")

# Model3
predict_3 <- predict(TARGET_FLAG_m3,newdata=testing_1,type='response')
glm.pred3 = ifelse(predict_3 > 0.5, 1, 0)
cM3 <- confusionMatrix(glm.pred3, testing_1$TARGET_FLAG, positive = "1")
```

```{r echo=FALSE}
df1b <- as.data.frame(cM1$byClass)
df1a <- as.data.frame(cM1$overall)
colnames(df1a) <- 'Model1'
colnames(df1b) <- 'Model1'
df1 <- rbind(df1a, df1b)

df2b <- as.data.frame(cM2$byClass)
df2a <- as.data.frame(cM2$overall)
colnames(df2a) <- 'Model2'
colnames(df2b) <- 'Model2'
df2 <- rbind(df2a, df2b)

df3b <- as.data.frame(cM3$byClass)
df3a <- as.data.frame(cM3$overall)
colnames(df3a) <- 'Model3'
colnames(df3b) <- 'Model3'
df3 <- rbind(df3a, df3b)

df <- cbind(df1,df2,df3)
kable(df)
```

### ROC Curve and Area Under the Curve

```{r}
rc1 <- roc(factor(TARGET_FLAG) ~ predict_1, data=testing_1)
rc2 <- roc(factor(TARGET_FLAG) ~ predict_2, data=testing_1)
rc3 <- roc(factor(TARGET_FLAG) ~ predict_3, data=testing_1)

plot(rc1,main='Model 1 - ROC Curve')
plot(rc2,main='Model 2 - ROC Curve')
plot(rc3,main='Model 3 - ROC Curve')


model <- c('Model 1', 'Model 2', 'Model 3')
area <- c(auc(rc1),auc(rc2),auc(rc3))
df <- data.frame(Model=model,AUC=area)
kable(df,caption='Area under the curve')
```

### Log-likelihood/AIC/BIC

```{r}
# Log-likelihood
LL.1 <- logLik(TARGET_FLAG_m1)
LL.2 <- logLik(TARGET_FLAG_m2)
LL.3 <- logLik(TARGET_FLAG_m3)
LL <- rbind(LL.1, LL.2, LL.3) %>% round(2)
```

```{r}
# Akaike Information Criterion
AIC.1 <- AIC(TARGET_FLAG_m1)
AIC.2 <- AIC(TARGET_FLAG_m2)
AIC.3 <- AIC(TARGET_FLAG_m3)
AIC <- rbind(AIC.1, AIC.2, AIC.3) %>% round(2)
```

```{r}
# BIC
BIC.1 <- BIC(TARGET_FLAG_m1)
BIC.2 <- BIC(TARGET_FLAG_m2)
BIC.3 <- BIC(TARGET_FLAG_m3)
BIC <- rbind(BIC.1, BIC.2, BIC.3) %>% round(2)
```

```{r}
eval.table <- cbind(LL, AIC, BIC)

rownames(eval.table) <- c("Model 1", "Model 2", "Model 3")
colnames(eval.table) <- c("Log Likelihood", "AIC", "BIC")

kable(eval.table)
```

### Checking variance inflation factors
```{r}
V1 <- vif(TARGET_FLAG_m1)
V2 <- vif(TARGET_FLAG_m2)
V3 <- vif(TARGET_FLAG_m3)
V1; V2; V3
```

## TARGET_AMT
### Key model statistics results
```{r}
col_mdl_names <- c("Model 1", "Model 2", "Model 3")

#Calculate mean squared errors for each model
mse <- function(sm) 
    mean(sm$residuals^2)

if(FALSE){
  cat("Mean Squared Error of Model 1:")
  mse(TARGET_AMT_m1)
  cat("Mean Squared Error of Model 2:")
  mse(TARGET_AMT_m2)
  cat("Mean Squared Error of Model 3:")
  mse(TARGET_AMT_m3)
}

col_mse <- c(mse(TARGET_AMT_m1),mse(TARGET_AMT_m2),mse(TARGET_AMT_m3))

# Calculate R^2 for each model:
if(FALSE){
  cat("R Squared of Model 1:")
  summary(TARGET_AMT_m1)$r.squared 
  cat("R Squared of Model 2:")
  summary(TARGET_AMT_m2)$r.squared 
  cat("R Squared of Model 3")
  summary(TARGET_AMT_m3)$r.squared 
}

col_r_sq <- c(summary(TARGET_AMT_m1)$r.squared, summary(TARGET_AMT_m2)$r.squared, summary(TARGET_AMT_m3)$r.squared)

if(FALSE){
  cat("F-Stat of Model 1:")
  summary(aov(TARGET_AMT_m1))[[1]]$F[1]
  cat("F-Stat of Model 2:")
  summary(aov(TARGET_AMT_m2))[[1]]$F[1]
  cat("F-Stat of Model 3")
  summary(aov(TARGET_AMT_m1))[[1]]$F[1]
}

col_f_stat <- c(summary(aov(TARGET_AMT_m1))[[1]]$F[1], summary(aov(TARGET_AMT_m2))[[1]]$F[1], summary(aov(TARGET_AMT_m3))[[1]]$F[1])
summary_df <- data.frame(cbind(col_mdl_names, col_mse, col_r_sq, col_f_stat))

colnames(summary_df) <- c("Model Name", "Mean Sq. Error", "R Squared", "F Stat")
kable(summary_df)
```

### Residual plots for each model

```{r}
# Model 1
qqnorm(TARGET_AMT_m1$residuals)
qqline(TARGET_AMT_m1$residuals)
INDEX <- seq(1,length(training_2$TARGET_AMT))
plot(TARGET_AMT_m1$residuals ~ INDEX,
     xlab='',
     ylab='Residuals',
     main='Residual Plot of Model 1')
abline(h=0,lty=3)
hist(TARGET_AMT_m1$residuals)

# Model 2
qqnorm(TARGET_AMT_m2$residuals)
qqline(TARGET_AMT_m2$residuals)
INDEX <- seq(1,length(training_2$TARGET_AMT))
plot(TARGET_AMT_m2$residuals ~ INDEX,
     xlab='',
     ylab='Residuals',
     main='Residual Plot of Model 2')
abline(h=0,lty=3)
hist(TARGET_AMT_m2$residuals)

# Model 3
qqnorm(TARGET_AMT_m3$residuals)
qqline(TARGET_AMT_m3$residuals)
INDEX <- seq(1,length(training_2$TARGET_AMT))
plot(TARGET_AMT_m3$residuals ~ INDEX,
     xlab='',
     ylab='Residuals',
     main='Residual Plot of Model 3')
abline(h=0,lty=3)
hist(TARGET_AMT_m3$residuals)
```

### Using testing data to check the predicting result
```{r}
actual <- testing_2$TARGET_AMT
new_data = select(testing_2, -TARGET_AMT)
predicted1 <- predict(TARGET_AMT_m1,newdata=new_data)
predicted2 <- predict(TARGET_AMT_m2,newdata=new_data)
predicted3 <- predict(TARGET_AMT_m3,newdata=new_data)

INDEX <- seq(1,length(testing_2$TARGET_AMT))
result1 <- data.frame(INDEX=INDEX, actual=actual, predicted1=predicted1)
result2 <- data.frame(INDEX=INDEX, actual=actual, predicted2=predicted2)
result3 <- data.frame(INDEX=INDEX, actual=actual, predicted3=predicted3)
result1 <- gather(result1,Group,TARGET_AMT,2:3)
result2 <- gather(result2,Group,TARGET_AMT,2:3)
result3 <- gather(result3,Group,TARGET_AMT,2:3)
p1 <- ggplot(data=result1,aes(INDEX,TARGET_AMT)) + 
  geom_point(aes(colour = factor(Group))) + ggtitle('Model 1')
p2 <- ggplot(data=result2,aes(INDEX,TARGET_AMT)) + 
  geom_point(aes(colour = factor(Group))) + ggtitle('Model 2')
p3 <- ggplot(data=result3,aes(INDEX,TARGET_AMT)) + 
  geom_point(aes(colour = factor(Group))) + ggtitle('Model 3')
p1; p2; p3
```

Overall, for predicting `TARGET_FLAG` Model 2 handled **Multicollinearity** issuse appropriately. Meanwhile, Model 2 slightly stands out after considering Accuracy, Sensitivity, Specificity, AUC, Log-likelihood number, AIC and BIC. 

For predicting `TARGET_AMT`, all of three models are not good enough in this case. More advanced modeling techniques might be helpful, although intution tells me that it would be very difficult to build a good model to predict `TARGET_AMT`. How much the cost would be in car accident is a highly random event.